---
title: "MultiOcc_Anal"
author: "Danielle Ethier"
date: '2022-07-06'
output: html_document
editor_options: 
  chunk_output_type: console
---

Now that we have build the multi-species occupancy model framework, we can test the analysis. 

#Load libraries

```{r libraries}

#You can install naturecounts from GitHub with the remotes package if you don't have it already. 
#install.packages("remotes")
#remotes::install_github("BirdsCanada/naturecounts")

require(naturecounts)
require(tidyverse)
require(unmarked)
require(lubridate)
require(reshape)
require(reshape2)
require(nloptr)
library(beepr)
require(ggpubr)
library(DMwR2)

out.dir <- paste("Output/")

```

#Format the data for the unmarked framework

```{r unmarked}

#species data
BAEA<-read.csv("BAEAdetect.csv", row.names="X")
SUSC<-read.csv("SUSCdetect.csv", row.names="X")
WWSC<-read.csv("WWSCdetect.csv", row.names="X")
BLSC<-read.csv("BLSCdetect.csv", row.names="X")
LTDU<-read.csv("LTDUdetect.csv", row.names="X")

y_list<-list(BAEA=as.matrix(BAEA), 
             SUSC=as.matrix(SUSC), 
             WWSC=as.matrix(WWSC), 
             BLSC=as.matrix(BLSC), 
             LTDU=as.matrix(LTDU))

#Detection level covariates 
doy<-read.csv("det_doy.csv", row.name="X")
tod<-read.csv("det_tod.csv", row.name="X")
vis<-read.csv("det_vis.csv", row.name="X")
dur<-read.csv("det_dur.csv", row.name="X")

det_list<-list(TimeofDay=tod, DayofYear = doy, Visibility=vis, Duration = dur)

#Observation level covaraites
occ_covs<-read.csv("obsCovs.csv", row.names="X")

yr.min<-2000
yr.max<-2022

mean<-2011.2349132948
sd<-6.65745963217531

```

#Fit a model with no covariates, assuming independence

To start, we assume independence among species. We do this by only allowing 1st order natural parameters (maxOrder = 1). This is the same as fitting 3 single-species occupancy models

```{r fit1}

msom_data<-unmarkedFrameOccuMulti(y=y_list, siteCovs=occ_covs, obsCovs=det_list)

#msom_data<-unmarkedFrameOccuMulti(y=y_list, siteCovs=occ_covs)

#fit model with no covariates to test 
summary(msom_data) # to get summary of dataset. Useful output for results of MS. 

fit_1 <- occuMulti(detformulas = c('~1', '~1', '~1', '~1', '~1'),
                   stateformulas = c('~1', '~1', '~1', '~1', '~1'),
                   maxOrder = 1,
                   data = msom_data)
summary(fit_1)

#AIC = 41449 
```

#Fit a model with no covarites (intercept-only), assuming dependence

```{r fit2}

fit_2<- occuMulti(detformulas = c('~1', '~1', '~1', '~1', '~1'),
                   stateformulas = c('~1', '~1', '~1', '~1', '~1','~1', '~1', '~1', '~1', '~1','~1', '~1', '~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_2)

#AIC = 39538, suggesting that there is strong support for the model incorporating dependence.

#BAEA occur together more frequently than expected by chance with all the associated prey, expect LTDU.
```

#Fit a model covarites for maginal model

Detection covaraites fixed as visibility and duration (based on de Zwann)

Interested in the effect of year on occupancy probability, while taking into account spatial autocorrelation using lat and long. 

```{r fit3}

fit_M1<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration', '~ Visibility + Duration', '~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+latitude*longitude', '~year_st+latitude*longitude', '~year_st+latitude*longitude', '~year_st+latitude*longitude', '~year_st+latitude*longitude', '~1', '~1', '~1', '~1', '~1','~1', '~1', '~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_M1)

beep(4)

#AIC = 38722

```


#M1:Marginal occupancy probability over time, latitude, and longitude

Omit the `cond` argument if you wish to calculate marginal occupancy probability
Averages over presence / absence of all other species


```{r predict marginal}

sp.list<-c("BAEA", "SUSC", "WWSC", "BLSC", "LTDU")

#Margial effect of year  
nd_marg <- data.frame(
 year_st = seq(min(occ_covs$year_st), max(occ_covs$year_st),length.out = 100),
 latitude = rep(mean(occ_covs$latitude), 100), 
 longitude=rep(mean(occ_covs$longitude), 100)
)

#sp loop
for(m in 1:length(sp.list)) {
  
 # m<-1 #for testing
  
marg <- predict(fit_M1, type = 'state', species = sp.list[m], newdata = nd_marg)

gg_df_marg <- data.frame(
  hd = nd_marg$year_st,
  occupancy = marg$Predicted,
  low = marg$lower,
  high = marg$upper
)

#Revert year_st back to Year
gg_df_marg<-gg_df_marg %>% mutate(Year=(hd+1)+yr.min)   

marg_fig <- ggplot(gg_df_marg, aes(x = Year, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('Marginal occupancy probability') +
  xlab('Year') +
  theme(text = element_text(size = 25))+
  theme_classic()

marg_fig

ggsave(paste(out.dir, sp.list[m], "_marginal_time.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

} #end sp Year loop

beep(4)
```


#Conditional occupancy probability

Calculation of conditional and marginal occupancy probabilities is done with the predict function.

Create a data.frame for predictions
The procedure is equivalent to creating data frames for all other applications of predict

Include complete range of observed *year_st*; hold all other variables at their mean. 

Predicting prey species occurrence when bald eagles are present

`Species` indicates which species we assume when predicting occupancy
`cond` indicates which species we are assuming is present or absent


```{r predict year}

#sp list excluding BAEA
sp.list<-c("SUSC", "WWSC", "BLSC", "LTDU")

yr_cond <- data.frame(
  latitude = rep(mean(occ_covs$latitude), 100),
  longitude = rep(mean(occ_covs$longitude), 100),
  year_st = seq(min(occ_covs$year_st), max(occ_covs$year_st),length.out = 100))

#sp loop
for(n in 1:length(sp.list)) {
  
BAEA_mod <- predict(fit_M1, type = 'state', species = sp.list[n], cond = 'BAEA', newdata = yr_cond)

#putting a `-`in front of BAEA tells predict you wish to assume they are absent

BAEA_mod_0 <- predict(fit_M1, type = 'state', species = sp.list[n], cond = '-BAEA', newdata = yr_cond)

#formatting ggplot

gg_df_cond <- data.frame(
  year_st = rep(yr_cond$year_st, 2),
  occupancy = c(BAEA_mod$Predicted,
                BAEA_mod_0$Predicted),
  low = c(BAEA_mod$lower,
          BAEA_mod_0$lower),
  high = c(BAEA_mod$upper,
           BAEA_mod_0$upper),
  conditional = rep(c('BAEA Present', 'BAEA Absent'),each = 100)
)

#Revert year_st back to Year
gg_df_cond<-gg_df_cond %>% mutate(Year=(year_st+1)+yr.min)

cond_fig <- ggplot(gg_df_cond, aes(x = Year, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional)) +
  geom_line() +
  ylab('Conditionaloccupancy probability') +
  xlab('Year') +
  labs(fill = '') +
  theme(text = element_text(size = 25),
        legend.position = c(0.75, 0.85))+
  theme_classic()

cond_fig

ggsave(paste(out.dir, sp.list[n], "_conditional_time.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

} # end sp con loop

beep(4)

```

```{r predict latitude}

lat_cond <- data.frame(
  year_st = rep(mean(occ_covs$year_st), 100),
  longitude = rep(mean(occ_covs$longitude), 100),
  latitude = seq(min(occ_covs$latitude), max(occ_covs$latitude),length.out = 100))

BAEA_SUSC <- predict(fit_M1, type = 'state', species = 'SUSC', cond = 'BAEA', newdata = lat_cond)

#putting a `-`in front of BAEA tells predict you wish to assume they are absent

BAEA_SUSC_0 <- predict(fit_M1, type = 'state', species = 'SUSC', cond = '-BAEA', newdata = lat_cond)

#formatting ggplot

gg_df_cond <- data.frame(
  latitude = rep(lat_cond$latitude, 2),
  occupancy = c(BAEA_SUSC$Predicted,
                BAEA_SUSC_0$Predicted),
  low = c(BAEA_SUSC$lower,
          BAEA_SUSC_0$lower),
  high = c(BAEA_SUSC$upper,
           BAEA_SUSC_0$upper),
  conditional = rep(c('BAEA Present', 'BAEA Absent'),each = 100)
)

cond_fig <- ggplot(gg_df_cond, aes(x = latitude, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional)) +
  geom_line() +
  ylab('Conditional SUSC\noccupancy probability') +
  xlab('Latitude') +
  labs(fill = '') +
  theme(text = element_text(size = 25),
        legend.position = c(0.75, 0.85))+
  theme_classic()

cond_fig


```

