---
title: "MultiOcc_Anal"
author: "Danielle Ethier"
date: '2022-07-06'
output: html_document
editor_options: 
  chunk_output_type: console
---

Now that we have build the multi-species occupancy model framework, we can test the analysis. 

##Review suggested an analysis with all seaduck pooled together. This is the requested analysis. 

#Load libraries

```{r libraries}

#You can install naturecounts from GitHub with the remotes package if you don't have it already. 
#install.packages("remotes")
#remotes::install_github("BirdsCanada/naturecounts")

require(naturecounts)
require(tidyverse)
require(unmarked)
require(lubridate)
require(reshape)
require(reshape2)
require(nloptr)
library(beepr)
require(ggpubr)
library(DMwR2)

#library(rgdal)
library(sf)
library(ggmap)
library(leaflet)
library(mapview)
library(rnaturalearth)


library("ggmap") 
register_stadiamaps("1cdeb586-3f5a-4ef9-813b-02375efa9f21") 


out.dir <- paste("Output/")

```

#Format the data for the unmarked framework

```{r unmarked}

#species data BCCWS
BAEA1<-read.csv("BAEAdetectBCCWS2.csv", row.names="X")
ALL1<-read.csv("AlldetectBCCWS2.csv", row.names="X")

#Detection level covariates BCCWS
doy1<-read.csv("detBCCWS_doy.csv", row.name="X")
tod1<-read.csv("detBCCWS_tod.csv", row.name="X")
vis1<-read.csv("detBCCWS_vis.csv", row.name="X")
dur1<-read.csv("detBCCWS_dur.csv", row.name="X")

#species data PSSS
BAEA2<-read.csv("BAEAdetectPSSS2.csv", row.names="X")
ALL2<-read.csv("AlldetectPSSS2.csv", row.names="X")

#Detection level covariates PSSS
doy2<-read.csv("detPSSS_doy.csv", row.name="X")
tod2<-read.csv("detPSSS_tod.csv", row.name="X")
vis2<-read.csv("detPSSS_vis.csv", row.name="X")
dur2<-read.csv("detPSSS_dur.csv", row.name="X")

BAEA<-rbind(BAEA1, BAEA2)
ALL<-rbind(ALL1, ALL2)

y_list<-list(BAEA=as.matrix(BAEA), 
             ALL=as.matrix(ALL))

doy<-rbind(doy1, doy2)
tod<-rbind(tod1, tod2)
vis<-rbind(vis1,vis2)
dur<-rbind(dur1, dur2)

det_list<-list(TimeofDay=tod, DayofYear = doy, Visibility=vis, Duration = dur)

#Observation level covaraites
occBCCWS_covs<-read.csv("obsCovsBCCWS.csv", row.names="X")
occPSSS_covs<-read.csv("obsCovsPSSS.csv", row.names="X")

occ_covs<-rbind(occBCCWS_covs, occPSSS_covs)

plot_data<-occ_covs %>% select(SiteCode, period, latitude, longitude)

#Standardize Lat but first get values you need to back transform later
latmin = min(occ_covs$latitude)
latmax = max(occ_covs$latitude)
latmean = mean(occ_covs$latitude)
latsd = sd(occ_covs$latitude)

#Standardize Lat but first get values you need to back transform later
longmin = min(occ_covs$longitude)
longmax = max(occ_covs$longitude)
longmean = mean(occ_covs$longitude)
longsd = sd(occ_covs$longitude)

lat.s<-scale(occ_covs$latitude)
long.s<-scale(occ_covs$longitude)

occ_covs<-occ_covs %>% mutate_at(c('latitude', 'longitude'), ~(scale(.) %>% as.vector)) %>% select(year_st, latitude, longitude)

yr.min<-2009
yr.max<-2022

```

Make a map
```{r plot}

#Create spatial data
plot <- plot_data %>% select("longitude", "latitude")%>% distinct()
plot_sf <- st_as_sf(plot, coords = c("longitude", "latitude"), crs = 4326)

#Create the base layer map
map <- get_stadiamap(bbox = as.numeric(st_bbox(plot_sf)), zoom = 5)

#Create a new plot for each year
plot<-ggmap(map) + 
  geom_point(data = plot_data, aes(x = longitude, y = latitude))
print(plot)

```

#Create the occupancy model framework and get summary outputs

```{r occ sum}

msom_data<-unmarkedFrameOccuMulti(y=y_list, siteCovs=occ_covs, obsCovs=det_list)

summary(msom_data)

plot(msom_data)

```


#Fit a model with no covariates, assuming independence

To start, we assume independence among species. We do this by only allowing 1st order natural parameters (maxOrder = 1). This is the same as fitting 5 single-species occupancy models

```{r fit1}

#fit model with no covariates to test 
summary(msom_data) # to get summary of dataset. Useful output for results of MS. 

fit_1 <- occuMulti(detformulas = c('~1', '~1'),
                   stateformulas = c('~1', '~1'),
                   maxOrder = 1,
                   data = msom_data)
summary(fit_1)

#AIC = 21995


```

#Fit a model with no covarites (intercept-only), assuming dependence

```{r fit2}

fit_2<- occuMulti(detformulas = c('~1', '~1'),
                   stateformulas = c('~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_2)

#AIC = 21992, suggesting that there is some support for the model incorporating dependence.

```

#Fit a model covarites for duration only
 
```{r fitD}

fit_D<- occuMulti(detformulas = c('~ Duration', '~ Duration'),
                   stateformulas = c('~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_D)

#AIC = 21972 

```

#Fit a model covarites for visibility

```{r fitV}

fit_V<- occuMulti(detformulas = c('~ Visibility', '~ Visibility'),
                   stateformulas = c('~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_V)

#AIC = 21983


```

#Fit a model covarites for duration and visibility
**Top Detection Model**

```{r fitDV}

fit_DV<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_DV)

#AIC = 21964 

```


#Fit a model covarites for maginal model (year only)


```{r fit year}

fit_M0<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~year_st'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_M0)


#AIC = 21933

fit_M02<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_M02)

#AIC = 21932


```

#Fit a model covarites for maginal model (with year+lat and lat)


```{r fit lat}

fit_lat<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+latitude', '~year_st+latitude', '~year_st+latitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_lat)

#AIC = 21445

fit_lat0<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~latitude', '~latitude', '~latitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_lat0)

#AIC: 21478

fit_lat1<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+latitude', '~year_st+latitude', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_lat1)

#AIC = 21497

fit_lat01<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~latitude', '~latitude', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_lat01)

#AIC: 21528

```


#Fit a model covarites for maginal model (with year+long and long)

```{r fit long}

fit_long<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+longitude', '~year_st+longitude', '~year_st+longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_long)

#AIC = 21527

#get some summary outputs

fit_long0<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~longitude', '~longitude', '~longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_long0)

#AIC: 21562


fit_long1<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+longitude', '~year_st+longitude', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_long1)

#AIC = 21583

fit_long01<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~longitude', '~longitude', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_long01)

#AIC: 21620

```

#Fit a model full model 
**Best Model**

Detection covaraites fixed as visibility and duration (based on de Zwann)

Interested in the effect of year on occupancy probability, while taking into account spatial autocorrelation using lat and long. 

```{r fit4}

#AIC = 21210

fit_M11<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+latitude+longitude', '~year_st+latitude+longitude', '~1'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_M11)

#AIC = 21236

#get some summary outputs of top model

Out_det<-summary(fit_M11)$det
write.csv(Out_det, "Output/Out_det2.csv")
Out_state<-summary(fit_M11)$state
write.csv(Out_state, "Output/Out_state2.csv")

```

#Fit second order models (with year and lat and long)

             
```{r fit test}

fit_Y<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~year_st'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_Y)

fit_lat<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~latitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_lat)

fit_long<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_long)

fit_latlong<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~latitude +longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_latlong)

fit_latlongyear<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~latitude +longitude+year_st'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_latlongyear)

fit_YY<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~year_st'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_YY)

fit_latY<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~latitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_latY)

fit_longY<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_longY)

fit_latlongY<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~latitude +longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_latlongY)


fit_M1<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st', '~year_st', '~year_st+latitude+longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_M1)

fit_FULL<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~year_st+latitude+longitude', '~year_st+latitude+longitude', '~year_st+latitude+longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_FULL)



fit_LLY<- occuMulti(detformulas = c('~ Visibility + Duration', '~ Visibility + Duration'),
                   stateformulas = c('~1', '~1', '~year_st+latitude+longitude'),
                   maxOrder = 2,
                   data = msom_data)
summary(fit_LLY)

```


#Model Selection
```{r model selection}

mods1 <- fitList(fit_1, fit_2)
modSel(mods1)
                
mods2 <- fitList(fit_D, fit_V, fit_DV) 
modSel(mods2)
                
mods3 <- fitList(fit_M0, fit_M02, fit_lat0, fit_lat, fit_long0, fit_long, fit_M11)
modSel(mods3)   

mods4 <- fitList(fit_Y, fit_lat, fit_long, fit_latlong, fit_latlongyear, fit_YY, fit_latY, fit_longY, fit_latlongY, fit_M1, fit_FULL)
modSel(mods4)
```

**Top model is M11**

#M1:Marginal occupancy probability over time and latitude

Omit the `cond` argument if you wish to calculate marginal occupancy probability
Averages over presence / absence of all other species


```{r predict marginal time}

sp.list<-c("BAEA", "ALL")

time_marg2 <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 6, byrow = FALSE, dimnames = NULL))

names(time_marg2) <- c("hd", "occupancy", "low", "high", "Year", "Species")

#only need to create the table once per analysis   

write.table(time_marg2, file = paste(out.dir,  "time_marg2.csv"), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")
    

#Margial effect of year  
nd_marg <- data.frame(
 year_st = seq(min(occ_covs$year_st), max(occ_covs$year_st),length.out = 100),
 longitude = rep(mean(occ_covs$longitude), 100), 
 latitude = rep(mean(occ_covs$latitude), 100)
)

#sp loop
for(m in 1:length(sp.list)) {
  
 # m<-1 #for testing
  
marg <- predict(fit_M1, type = 'state', species = sp.list[m], newdata = nd_marg)

gg_df_marg <- data.frame(
  hd = nd_marg$year_st,
  occupancy = marg$Predicted,
  low = marg$lower,
  high = marg$upper
)

#Revert year_st back to Year
gg_df_marg<-gg_df_marg %>% mutate(Year=(hd-1)+yr.min)   
gg_df_marg$Species<-sp.list[m]

marg_fig <- ggplot(gg_df_marg, aes(x = Year, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('') +
  xlab('') +
  ggtitle(sp.list[m])+
  theme_classic()+
 # scale_fill_brewer(palette="Blues")+
  scale_x_continuous(limits = c(2010, 2022), breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(text = element_text(size = 30))   

marg_fig

ggsave(paste(out.dir, sp.list[m], "_marginal_time2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

write.table(gg_df_marg, file = paste(out.dir, "time_marg2.csv"), row.names = FALSE, append = TRUE, 	quote = FALSE, sep = ",", col.names = FALSE)

} #end sp Year loop

marg_time2<-read.csv("Output/ time_marg2.csv")
marg_time2<-na.omit(marg_time2)

ggplot(marg_time2, aes(x = Year, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('Marginal occupancy') +
  xlab('Year') +
  facet_wrap(~Species, scales="free")+
  theme_classic()+
  scale_x_continuous(limits = c(2010, 2022), breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(text = element_text(size = 20)) 


```


```{r predict marginal latitude}

sp.list<-c("BAEA", "ALL")

#Margial effect of lat
nd_marg <- data.frame(
  year_st = rep(mean(occ_covs$year_st), 100),
  longitude = rep(mean(occ_covs$longitude), 100),
  latitude = seq(min(occ_covs$latitude), max(occ_covs$latitude),length.out = 100))

lat_marg2 <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 6, byrow = FALSE, dimnames = NULL))

names(lat_marg2) <- c("hd", "occupancy", "low", "high",  "Lat", "Species")

#only need to create the table once per analysis   

write.table(lat_marg2, file = paste(out.dir,  "lat_marg2.csv"), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")
    

#sp loop
for(m in 1:length(sp.list)) {
  
 # m<-1 #for testing
  
marg <- predict(fit_M1, type = 'state', species = sp.list[m], newdata = nd_marg)

gg_df_marg <- data.frame(
  hd = nd_marg$latitude,
  occupancy = marg$Predicted,
  low = marg$lower,
  high = marg$upper
)

#Revert latitude back to natural scale 
#gg_df_marg<-gg_df_marg %>% mutate(Lat=hd*(latmax - latmin) + latmin)
gg_df_marg$Lat <- gg_df_marg$hd* attr(lat.s, 'scaled:scale') + attr(lat.s, 'scaled:center')

gg_df_marg$Species<-sp.list[m]

marg_fig <- ggplot(gg_df_marg, aes(x = Lat, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('') +
  xlab('') +
  ggtitle(sp.list[m])+
  theme_classic()+
  theme(text = element_text(size = 25))

marg_fig

ggsave(paste(out.dir, sp.list[m], "_marginal_latitude2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

write.table(gg_df_marg, file = paste(out.dir, "lat_marg2.csv"), row.names = FALSE, append = TRUE, 	quote = FALSE, sep = ",", col.names = FALSE)

} #end sp Year loop

marg_lat<-read.csv("Output/ lat_marg2.csv")
marg_lat<-na.omit(marg_lat)

ggplot(marg_lat, aes(x = Lat, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('Marginal occupancy') +
  xlab('Latitude') +
  facet_wrap(~Species, scales="free")+
  theme_classic()+
   theme(text = element_text(size = 20)) 

```


```{r predict marginal longitude}

sp.list<-c("BAEA", "ALL")

long_marg2 <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 6, byrow = FALSE, dimnames = NULL))

names(long_marg2) <- c("hd", "occupancy", "low", "high",  "Long", "Species")

#only need to create the table once per analysis   

write.table(long_marg2, file = paste(out.dir,  "long_marg2.csv"), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")

#Margial effect of long
nd_marg <- data.frame(
  year_st = rep(mean(occ_covs$year_st), 100),
  latitude = rep(mean(occ_covs$latitude), 100),
  longitude = seq(min(occ_covs$longitude), max(occ_covs$longitude),length.out = 100))

#sp loop
for(m in 1:length(sp.list)) {
  
 # m<-1 #for testing
  
marg <- predict(fit_M1, type = 'state', species = sp.list[m], newdata = nd_marg)

gg_df_marg <- data.frame(
  hd = nd_marg$longitude,
  occupancy = marg$Predicted,
  low = marg$lower,
  high = marg$upper
)

#Revert latitude back to natural scale 
#gg_df_marg<-gg_df_marg %>% mutate(Long=hd*(longmax - longmin) + longmin)

gg_df_marg$Long <- gg_df_marg$hd* attr(long.s, 'scaled:scale') + attr(long.s, 'scaled:center')

gg_df_marg$Species<-sp.list[m]

marg_fig <- ggplot(gg_df_marg, aes(x = Long, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('') +
  xlab('') +
  ggtitle(sp.list[m])+
  theme_classic()+
  theme(text = element_text(size = 25))

marg_fig

ggsave(paste(out.dir, sp.list[m], "_marginal_longitude2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

write.table(gg_df_marg, file = paste(out.dir, "long_marg2.csv"), row.names = FALSE, append = TRUE, 	quote = FALSE, sep = ",", col.names = FALSE)

} #end sp Year loop

marg_long<-read.csv("Output/ long_marg2.csv")
marg_long<-na.omit(marg_long)

ggplot(marg_long, aes(x = Long, y = occupancy)) +
  geom_ribbon(aes(ymin = low, ymax = high), alpha = 0.5) +
  geom_line() +
  ylab('Marginal occupancy') +
  xlab('Longitude') +
  facet_wrap(~Species, scales="free")+
  theme_classic()+
   scale_x_continuous(limits = c(-125.25, -122.25), breaks = c(-125, -124, -123, -122))+
   theme(text = element_text(size = 20)) 



```

#M1: Conditional occupancy probability

Calculation of conditional and marginal occupancy probabilities is done with the predict function.

Create a data.frame for predictions
The procedure is equivalent to creating data frames for all other applications of predict

Include complete range of observed *year_st*; hold all other variables at their mean. 
Predicting prey species occurrence when bald eagles are present

`Species` indicates which species we assume when predicting occupancy
`cond` indicates which species we are assuming is present or absent

#Year
```{r predict year}

yr_cond <- data.frame(
  latitude = rep(mean(occ_covs$latitude), 100),
  longitude = rep(mean(occ_covs$longitude), 100),
  year_st = seq(min(occ_covs$year_st), max(occ_covs$year_st),length.out = 100))

BAEA_mod <- predict(fit_M1, type = 'state', species = "ALL", cond = 'BAEA', newdata = yr_cond)

#putting a `-`in front of BAEA tells predict you wish to assume they are absent

BAEA_mod_0 <- predict(fit_M1, type = 'state', species = "ALL", cond = '-BAEA', newdata = yr_cond)

#formatting ggplot

gg_df_cond <- data.frame(
  year_st = rep(yr_cond$year_st, 2),
  occupancy = c(BAEA_mod$Predicted,
                BAEA_mod_0$Predicted),
  low = c(BAEA_mod$lower,
          BAEA_mod_0$lower),
  high = c(BAEA_mod$upper,
           BAEA_mod_0$upper),
  conditional = rep(c('BAEA Present', 'BAEA Absent'),each = 100)
)

#Revert year_st back to Year
gg_df_cond<-gg_df_cond %>% mutate(Year=(year_st+1)+yr.min)

cond_fig <- ggplot(gg_df_cond, aes(x = Year, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional)) +
  geom_line() +
  ylab('') +
  xlab('') +
  labs(fill = '') +
  theme_classic()+
  ggtitle(sp.list2[n])+
  scale_fill_brewer(palette="Blues")+
  scale_x_continuous(limits = c(2010, 2022), breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(text = element_text(size = 30)) 
#  theme(legend.position = "none")    
  
cond_fig

ggsave(paste(out.dir, sp.list2[n], "_conditional_time2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()


```


#Latitude
```{r predict latitude}

lat_cond <- data.frame(
  year_st = rep(mean(occ_covs$year_st), 100),
  longitude = rep(mean(occ_covs$longitude), 100),
  latitude = seq(min(occ_covs$latitude), max(occ_covs$latitude),length.out = 100))

BAEA_mod <- predict(fit_M1, type = 'state', species = "ALL", cond = 'BAEA', newdata = lat_cond)

#putting a `-`in front of BAEA tells predict you wish to assume they are absent

BAEA_mod_0 <- predict(fit_M1, type = 'state', species = "ALL", cond = '-BAEA', newdata = lat_cond)

#Revert latitude back to natural scale 
lat_cond<-lat_cond %>% mutate(Lat=latitude*(latmax - latmin) + latmin)

#formatting ggplot

gg_df_cond <- data.frame(
  latitude = rep(lat_cond$Lat, 2),
  occupancy = c(BAEA_mod$Predicted,
                BAEA_mod_0$Predicted),
  low = c(BAEA_mod$lower,
          BAEA_mod_0$lower),
  high = c(BAEA_mod$upper,
           BAEA_mod_0$upper),
  conditional = rep(c('BAEA Present', 'BAEA Absent'),each = 100)
)

cond_fig <- ggplot(gg_df_cond, aes(x = latitude, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional)) +
  geom_line() +
  ylab('') +
  xlab('') +
  labs(fill = '') +
   ggtitle(sp.list2[n])+
  scale_fill_brewer(palette="Blues")+
  theme_classic()+
  theme(text = element_text(size = 30))+
  theme(legend.position = "none")

cond_fig

ggsave(paste(out.dir, sp.list2[n], "_conditional_latitude2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

```


#Longitude
```{r predict latitude}

long_cond <- data.frame(
  year_st = rep(mean(occ_covs$year_st), 100),
  latitude = rep(mean(occ_covs$latitude), 100),
  longitude = seq(min(occ_covs$longitude), max(occ_covs$longitude),length.out = 100))

BAEA_mod <- predict(fit_M1, type = 'state', species = "ALL", cond = 'BAEA', newdata = long_cond)

#putting a `-`in front of BAEA tells predict you wish to assume they are absent

BAEA_mod_0 <- predict(fit_M1, type = 'state', species = "ALL", cond = '-BAEA', newdata = long_cond)

#Revert latitude back to natural scale 
long_cond<-long_cond %>% mutate(Long=longitude*(longmax - longmin) + longmin)

#formatting ggplot

gg_df_cond <- data.frame(
  longitude = rep(long_cond$Long, 2),
  occupancy = c(BAEA_mod$Predicted,
                BAEA_mod_0$Predicted),
  low = c(BAEA_mod$lower,
          BAEA_mod_0$lower),
  high = c(BAEA_mod$upper,
           BAEA_mod_0$upper),
  conditional = rep(c('BAEA Present', 'BAEA Absent'),each = 100)
)

cond_fig <- ggplot(gg_df_cond, aes(x = longitude, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional)) +
  geom_line() +
  ylab('') +
  xlab('') +
  labs(fill = '') +
   ggtitle(sp.list2[n])+
  scale_fill_brewer(palette="Blues")+
  theme_classic()+
  theme(text = element_text(size = 30))+
  theme(legend.position = "none")

cond_fig

ggsave(paste(out.dir, sp.list2[n], "_conditional_longitude2.jpeg", sep=""), width = 11, height = 8, device = "jpeg")
while (!is.null(dev.list()))  dev.off()

```

#Parameter Estimates

```{r parameters}

## Bootstrapping confidence intervals with 100 samples

#Detection
#Code incomplete, see 03
BAEA_det<-predict(fit_M11, 'det', species=1) #BAEA
ALL_det<-predict(fit_M11, 'det', species=2)#ALL

#Combine and plot
all_det<-rbind(BAEA_det[1,], ALL_det[1,])
all_det$Species<-c("BAEA", "ALL")

ggplot(data=all_det, aes(x=Species, y=Predicted))+
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.1) +
  geom_line()+
  geom_point()+
  theme_classic()+
  ylab("Detection probability") +
  theme(text = element_text(size = 20))

#Marginal occupancy
#Code incomplete, see 03
BAEA<-predict(fit_M1,'state',species=1) #BAEA
ALL<-predict(fit_M1,'state',species=2) #ALL


#Combine and plot
all_marg<-rbind(BAEA, ALL)
all_marg$Species<-c("BAEA", "ALL")

ggplot(data=all_marg, aes(x=Species, y=Predicted))+
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.1) +
  geom_line()+
  geom_point()+
  theme_classic()+
  ylab("Marginal occupancy probability") +
  xlab("")+
  theme(text = element_text(size = 20))
  
# conditional occupancy given bald eagles are present
ALL_P<-predict(fit_M1,'state',species=2,cond='BAEA') #ALL | BAEA
ALL_P<-ALL_P %>% mutate(mean= mean(Predicted), se=mean(SE), sd=sd(Predicted),  LCI=mean(lower), UCI = mean(upper))
ALL_P<-ALL_P %>% select(mean, se, LCI, UCI) %>% distinct()

# conditional occupancy given bald eagles are absent
ALL_A<-predict(fit_M1,'state',species=2,cond='-BAEA') #ALL | -BAEA
ALL_A<-ALL_A %>% mutate(mean= mean(Predicted), se=mean(SE), sd=sd(Predicted),  LCI=mean(lower), UCI = mean(upper))
ALL_A<-ALL_A %>% select(mean, se, LCI, UCI) %>% distinct()

#Combine and plot
all_cond2<-rbind(ALL_P, ALL_A)
all_cond2$Species<-c("ALL", "ALL")
all_cond2$BaldEagle<-c("Present", "Absent")

dodge <- position_dodge(width=0.5) 

ggplot(data=all_cond2, aes(x=Species, y=mean, color=BaldEagle))+
  geom_errorbar(aes(ymin=mean+se, ymax=mean-se), width=.1, position=dodge) +
  geom_line(position=dodge)+
  geom_point(position=dodge)+
  theme_classic()+
  labs(fill="Bald Eagle")+
  ylab("Conditional occupancy probability") +
  xlab("")+
  scale_colour_manual(values = c('black','darkgrey')) +
  theme(text = element_text(size = 20))


```

bt<-parboot((fit_M1, nsim=30))


unmarkedFitOccuMulti object describing the model fit.



