---
title: "DataExplore"
author: "Danielle Ethier"
date: '2022-06-25'
output: html_document
---

#Load libraries

```{r libraries}

#You can install naturecounts from GitHub with the remotes package if you don't have it already. 
#install.packages("remotes")
#remotes::install_github("BirdsCanada/naturecounts")

require(naturecounts)
require(tidyverse)

```

If you have not used the naturecount R package before to explore or access data, you may find this [Introductory Tutorial](https://birdscanada.github.io/NatureCounts_IntroTutorial/) helpful.  

#Explore the data on NatureCounts

Filtering based on geographic extent. Specifically, we will create a `bbox` (bounding box) around the Salish Sea to see what data are available in this area. Coordinates = c(left = -125.76937, bottom = 47.89851, right = -12242952, top = 51.09647) 

search_species_code("BAEA") #3020
search_species_code("SUSC") #770
search_species_code("WWSC") #45994
search_species_code("BLSC") #791
search_species_code("LTDU") #810

Winter data to correspond with the BCCWS (September 1 to April 30): c(244, 120)

Using the information above, we can query the database.

```{r explore}

BAEA<-nc_count(region = list(bbox = c(left = -125.76937, bottom = 47.89851, right = -122.42952, top = 51.09647)), species = c(3020), doy=c(244,120))
View(BAEA)

All<-nc_count(region = list(bbox = c(left = -125.76937, bottom = 47.89851, right = -122.42952, top = 51.09647)), species = c(3020, 770, 45994, 791, 810), doy=c(244,120))
View(All)
```

The most promising collections look like: BCCWS, CBC, CMMN-DET-RPBO, and EBIRD-CA-BC (may also want to look into EBIRD US, if within scope of the assessment)

RPBO (Rocky Point Bird Observatory): start year = 1998, Fall only 21 July - 18 Oct, Banding + Census, no longer collect waterbird specific data. 
